[Re] Marriage and happiness: Providing evidence against a relationship between inequality and happiness in Oishi, Kesebir, and Diener (2011)

### 2020-02-24 analysis/paper.Rmd
Fixed typos in paper.

@@

The dataset from gssr contained a larger time interval than the original paper. It was filtered back to match the original interval. Only years prior to 2008 are used in both analyses. This dataset was also merged with the original US Census Bureau data found in Gini_families.csv[@data] by appending the "total" column to the GSS data using "year" as the join key.

@@

Aside from gathering the input data from a new source, several different packages were used in addition to ensure the reproduced results were adequately tested. These will be discussed further in the respective sections below.

@@

Table 1 in the original publication explores the correlations between happiness and various control variables found in the dataset.

@@

The R function cor.test() was used to get the Pearson correlation coefficient between the different independent variables (income on a log scale, age, sex, white, married) and the response variable, happiness. Grundberg et al.'s code was slightly modified to keep the same encoding of variables given the new dataset input.

@@

We were able to almost exactly reproduce the results from Table 1 in the original paper that shown in Figure \@ref(fig:table1). The only difference is the estimate for Sex was 0.01 in the publication compared to 0.00 in our replication. This may be due to rounding versions or an underlying exclusion of outliers we saw with the above plot.

@@

Again we were able to almost identically reproduce the results from Table 2 in the original paper, shown in Figure \@ref(fig:table2). Aside from slight variability in the second and third decimal places, our results from the linear models are very close. There are several differences with the number of observations, listed in the bottom of the table, which is to be expected as we are using a package with the GSS data and it is likely not an identical snapshot to the one used in the original paper. However, this variability is minor in scale.Even though slight variabilities were observed in different cells throughout this table, no disparities were large enough to impact the significance of the results as denoted in the tables with asterisks. Therefore, even with new data sources we found that the conclusions hold for the linear model analysis.

@@

Under conditions of both a new source dataset and different R packages used for data manipulation and visualization, the 3 exhibits in the original publication were able to be replicated. With only minor discrepancies in the numbers themselves, the conclusions drawn from the paper's graphics were not significantly altered by any means. Successful replication of this paper is largely a result of transparency in the original publication providing clarity around data encoding and the models used for analysis. This provided a platform upon which we were able to adapt the code and generate results that in large part align directly with those in the publication.

>>>>

In the spirit of reproducible science, a replication of the paper "Marriage and happiness: Providing evidence against a relationship between inequality and happiness in Oishi, Kesebir, and Diener (2011)" [@marriagehappiness] is conducted according to the ReScience standards. ReScience defines replication as:

>"a scientific study (computational or other) means repeating a published protocol, respecting its spirit and intentions but varying the technical details. For computational work, this would mean using different software, running a simulation from different initial conditions, etc. The idea is to change something that everyone believes shouldn’t matter, and see if the scientific conclusions are affected or not" [@rescience].

This replication is unique because it is a replication of a replication. The first paper, "Income Inequality and Happiness" [@incomeinequality], is a study on happiness and specifically focuses on income equality. The study makes the conclusion that in years where there is greater income inequality in America, people are less happy. The paper attributes this negative relationship to meta variables such as perceived trust and perceived fairness in society. The second paper [@marriagehappiness] calls into question the data and the conclusions used by this study. The second paper replicates the first factoring in additional data like race, gender, and marriage to test for possible confounding variables. Most noteably, a relationship between marital status (single vs married) results in the only significant factor of a linear regression using all variables. In essence, the second paper makes the claim that when controlling for marriage, there is no relationship between income inequality and the happiness of Americans - invalidating the first paper's results.

This paper will now be a second replication, focused on critiquing the methods and traceability of the second paper. Throughout the analysis, some critical pieces of missing data led to challenges in reproducing the results. In an effort to reproduce results, we had to find our own data and sources to tie in for the analysis.

@@

The dataset from gssr contained a larger time interval than the original paper. It was filtered back to match the original interval so that only years prior to 2008 are used in both analyses. The new dataset was subsequently merged with the original US Census Bureau data found in Gini_families.csv[@data] by appending the "total" column to the GSS data using "year" as the join key.

@@

Aside from gathering the input data from a new source, several different packages were used to ensure the reproduced results were adequately tested. These will be discussed further in their respective sections below.

@@

Table 1 in the original publication explored the correlations between happiness and various control variables found in the dataset.

@@

The R function cor.test() was used to get the Pearson correlation coefficient between the different independent variables (income on a log scale, age, sex, race, married) and the response variable, happiness. Grundberg et al.'s code was slightly modified to keep the same encoding of variables given the new dataset input.

@@

We were able to almost exactly reproduce the results from Table 1 in the original paper as shown in Figure \@ref(fig:table1). The only difference being that the estimate for Sex was 0.01 in the publication compared to 0.00 in our replication. This may be due to rounding versions or an underlying exclusion of outliers we saw with the above plot.

@@

Again we were able to almost identically reproduce the results from Table 2 in the original paper, shown in Figure \@ref(fig:table2). Aside from slight variability in the second and third decimal places, our results from the linear models are very close. There are several differences with the number of observations, listed in the bottom of the table, which is to be expected as we are using a package with the GSS data and it is likely not an identical snapshot to the one used in the original paper. However, this variability is minor in scale. Even though slight variabilities were observed in different cells throughout this table, no disparities were large enough to impact the significance of the results as denoted in the tables with asterisks. Therefore, even with new data sources we found that the conclusions hold for the linear model analysis.

@@

Under conditions of both a new source dataset and different R packages used for data manipulation and visualization, the 3 exhibits in the original publication were able to be replicated. With only minor discrepancies in the numbers themselves, the conclusions drawn from the paper's graphics were not significantly altered. Successful replication of this paper is largely a result of transparency in the original publication providing clarity around data encoding and the models used for analysis. This provided a platform upon which we were able to adapt the code and generate results that in large part align directly with those in the publication.
